Think deep for reef resilience

### 2025-08-14 paper/paper_draft.qmd
Add repository citation to the manuscript

DHW values on a per-pixel basis at 1km^2^ resolution were then calculated as the 12-week
rolling sum of positive HotSpot SST Anomaly values that are greater than 1°C, expressed in
units of °C-weeks. The 12-week accumulation period reflects the cumulative nature of coral
thermal stress. Brief temperature spikes may be tolerable, but sustained or repeated
exposure to elevated temperatures causes progressively greater physiological stress and
bleaching severity. DHW values integrate both the magnitude and duration of thermal stress:
a moderate anomaly (e.g., 1°C) sustained for 12 weeks produces the same DHW value as a
severe anomaly (e.g., 3°C) lasting 4 weeks, though the biological impacts may differ. DHW
estimates for each indicative reef (see @GBRMPA2022) were then determined by taking the mean
DHW value for all pixels touching each reef outline. The process described here was
conducted for all 2429 reefs to create DHW projections spanning from 2025 to 2075. Code used
to calculate DHW data and perform analyses involved in this investigation can be found at
**\[insert repo citation\]**.

::: {#tbl-gcms}
| Global Climate Model source | ECS value (°C) | Interpreted relative severity | Results found in |
|-----------------------|-----------------------|------------------------|-----------------------|
| ACCESS-CM2 | 4.66 [@grose2023] | High | **insert repo citation** |
| EC-Earth3-Veg | 4.30 [@wyser2020] | High | **insert repo citation** |
| GFDL-CM4 | 4.1 [@sentman2025] | High | **insert repo citation** |
| ACCESS-ESM1-5 | 3.87 [@ziehn2020] | Intermediate | Main text |
| NorESM2-MM | 2.5 [@seland2020] | Low | **insert repo citation** |

@@

Timeseries clustering analyses were performed at the management area and reefal bioregion
scale (defined in GBRMPA [-@GBRMPA2024]). Additional analyses were also performed at the
whole-of-GBR scale. These were found to lead to identical conclusions and so are not shown
here. To compare timeseries clustering results across GCMs, management area scale clustering
results are used to preserve coarse scale latitudinal differences between regions, while
maintaining ease of comparison and visualisation. These comparisons include quantification
of differences between low and high cover clusters across management regions under each GCM
using the Complexity-Invariant Distance between cluster median timeseries. To assess how
different reef characteristics influence timeseries clustering at a finer scale, bioregion
results are used to capture patterns occurring within individual clusters of reefs. Results
of all analyses at all clustering scales can be found at \[**insert citation for
repository**\] for those interested.

@@

Each Global Climate Model (GCM) shows similar relationships between reef clusters and
contextual analysis factors, thus further comparisons of these clusters for each analysis
factor are only shown for ACCESS-ESM1-5 (an intermediate Equilibrium Climate Sensitivity
scenario [@sfig-ecs]). Results for the additional GCMs can be found at \[**BG: Insert
repository citation.**\]. However, there are substantial differences in the longevity of
reef resilience and maintenance of coral cover post 2030 between the GCMs, thus coral cover
timeseries are displayed for each GCM (displayed clustered at management-area scale to
preserve coarse spatial differences in cover trajectories) (@fig-gcm-cover-trajectories).

>>>>

DHW values on a per-pixel basis at 1km^2^ resolution were then calculated as the 12-week
rolling sum of positive HotSpot SST Anomaly values that are greater than 1°C, expressed in
units of °C-weeks. The 12-week accumulation period reflects the cumulative nature of coral
thermal stress. Brief temperature spikes may be tolerable, but sustained or repeated
exposure to elevated temperatures causes progressively greater physiological stress and
bleaching severity. DHW values integrate both the magnitude and duration of thermal stress:
a moderate anomaly (e.g., 1°C) sustained for 12 weeks produces the same DHW value as a
severe anomaly (e.g., 3°C) lasting 4 weeks, though the biological impacts may differ. DHW
estimates for each indicative reef (see @GBRMPA2022) were then determined by taking the mean
DHW value for all pixels touching each reef outline. The process described here was
conducted for all 2429 reefs to create DHW projections spanning from 2025 to 2075. Code used
to calculate DHW data and perform analyses involved in this investigation can be found at
Grier and Iwanaga [-@grier2025].

::: {#tbl-gcms}
| Global Climate Model source | ECS value (°C)     | Interpreted relative severity | Results found in |
|-----------------------|-----------------------|------------------------|-----------------------|
| ACCESS-CM2                  | 4.66 [@grose2023]  | High                          | [@grier2025]     |
| EC-Earth3-Veg               | 4.30 [@wyser2020]  | High                          | [@grier2025]     |
| GFDL-CM4                    | 4.1 [@sentman2025] | High                          | [@grier2025]     |
| ACCESS-ESM1-5               | 3.87 [@ziehn2020]  | Intermediate                  | Main text        |
| NorESM2-MM                  | 2.5 [@seland2020]  | Low                           | [@grier2025]     |

@@

Timeseries clustering analyses were performed at the management area and reefal bioregion
scale (defined in GBRMPA [-@GBRMPA2024]). Additional analyses were also performed at the
whole-of-GBR scale. These were found to lead to identical conclusions and so are not shown
here. To compare timeseries clustering results across GCMs, management area scale clustering
results are used to preserve coarse scale latitudinal differences between regions, while
maintaining ease of comparison and visualisation. These comparisons include quantification
of differences between low and high cover clusters across management regions under each GCM
using the Complexity-Invariant Distance between cluster median timeseries. To assess how
different reef characteristics influence timeseries clustering at a finer scale, bioregion
results are used to capture patterns occurring within individual clusters of reefs. Results
of all analyses at all clustering scales can be found at Grier and Iwanaga [-@grier2025] for
those interested.

@@

Each Global Climate Model (GCM) shows similar relationships between reef clusters and
contextual analysis factors, thus further comparisons of these clusters for each analysis
factor are only shown for ACCESS-ESM1-5 (an intermediate Equilibrium Climate Sensitivity
scenario [@sfig-ecs]). Results for the additional GCMs can be found at Grier and Iwanaga
[-@grier2025]. However, there are substantial differences in the longevity of reef
resilience and maintenance of coral cover post 2030 between the GCMs, thus coral cover
timeseries are displayed for each GCM (displayed clustered at management-area scale to
preserve coarse spatial differences in cover trajectories) (@fig-gcm-cover-trajectories).
